### 2022-03-07 paper/paper.tex
WIP on laptop

% This is an example of using latex for a paper/report of specified
% size/layout. It's useful if you want to provide a PDF that looks
% like it was made in a normal word processor.

% While writing, don't stop for errors
\nonstopmode

% Use the article doc class, with an 11 pt basic font size
%\documentclass[11pt, a4paper, draft]{article}
\documentclass[11pt, a4paper]{article}

% Makes the main font Nimbus Roman, a Times New Roman lookalike:
%\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx
% OR use this for proper Times New Roman (from msttcorefonts package
% on Ubuntu). Use xelatex instead of pdflatex to compile:
\usepackage{fontspec}
\usepackage{xltxtra}
\usepackage{xunicode}
\defaultfontfeatures{Scale=MatchLowercase,Mapping=tex-text}
\setmainfont{Times New Roman}

% Nice citations
\usepackage{natbib}

% Set margins
\usepackage[margin=2.5cm]{geometry}

% Multilingual support
\usepackage[english]{babel}

% Nice mathematics
\usepackage{amsmath}

% Left right harpoons for kinetic equations
\usepackage{mathtools}

% Control over maketitle
\usepackage{titling}

% Section styling
\usepackage{titlesec}

% Ability to use colour in text
\usepackage[usenames]{color}
\definecolor{grey}{rgb}{0.6, 0.6, 0.63}
\definecolor{black}{rgb}{0, 0, 0}

% For the \degree symbol
\usepackage{gensymb}

% Allow includegraphics and nice wrapped figures
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[outercaption]{sidecap}
\usepackage{subfigure}

% Nice quotes
\usepackage{csquotes}

% Set formats using titlesec
\titleformat*{\section}{\bfseries\rmfamily}
\titleformat*{\subsection}{\bfseries\itshape\rmfamily}

% thetitle is the number of the section. This sets the distance from
% the number to the section text.
\titlelabel{\thetitle.\hskip0.3em\relax}

% Set title spacing with titlesec, too.  The first {1.0ex plus .2ex
% minus .7ex} sets the spacing above the section title. The second
% {-1.0ex plus 0.2ex} sets the spacing the section title to the
% paragraph.
\titlespacing{\section}{0pc}{1.0ex plus .2ex minus .7ex}{-1.1ex plus 0.2ex}

%% Trick to define a language alias and permit language = {en} in the .bib file.
% From: http://tex.stackexchange.com/questions/199254/babel-define-language-synonym
\usepackage{letltxmacro}
\LetLtxMacro{\ORIGselectlanguage}{\selectlanguage}
\makeatletter
\DeclareRobustCommand{\selectlanguage}[1]{%
  \@ifundefined{alias@\string#1}
    {\ORIGselectlanguage{#1}}
    {\begingroup\edef\x{\endgroup
       \noexpand\ORIGselectlanguage{\@nameuse{alias@#1}}}\x}%
}
\newcommand{\definelanguagealias}[2]{%
  \@namedef{alias@#1}{#2}%
}
\makeatother
\definelanguagealias{en}{english}
\definelanguagealias{eng}{english}
%% End language alias trick

%% Any aliases here
\newcommand{\mb}[1]{\mathbf{#1}} % this won't work?
% Emphasis and bold.
\newcommand{\e}{\emph}
\newcommand{\code}[1]{\textsf{#1}}
\newcommand{\dvrg}{\nabla\vcdot\nabla}
%% END aliases

% Custom font defs
% fontsize is \fontsize{fontsize}{linespacesize}
\def\authorListFont{\fontsize{11}{11} }
\def\corrAuthorFont{\fontsize{10}{10} }
\def\affiliationListFont{\fontsize{11}{11}\itshape }
\def\titleFont{\fontsize{14}{11} \bfseries }
\def\textFont{\fontsize{11}{11} }
\def\sectionHdrFont{\fontsize{11}{11}\bfseries}
\def\bibFont{\fontsize{10}{10} }
\def\captionFont{\fontsize{10}{10} }

% Caption font size to be small.
\usepackage[font=small,labelfont=bf]{caption}

% Make a dot for the dot product, call it vcdot for 'vector calculus
% dot'. Bigger than \cdot, smaller than \bullet.
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

\def\firstAuthorLast{James}

% Affiliations
\def\Address{\\
\affiliationListFont Adaptive Behaviour Research Group, Department of Psychology,
  The University of Sheffield, Sheffield, UK \\
}

% The Corresponding Author should be marked with an asterisk. Provide
% the exact contact address (this time including street name and city
% zip code) and email of the corresponding author
\def\corrAuthor{Seb James}
\def\corrAddress{Department of Psychology, The University of Sheffield,
  Western Bank, Sheffield, S10 2TP, UK}
\def\corrEmail{seb.james@sheffield.ac.uk}

% Figure out the font for the author list..
\def\Authors{\authorListFont Sebastian~S.~James, Stuart~P.~Wilson  \Address \\
  \corrAuthorFont $^{*}$ Correspondence: \corrEmail}

% No page numbering please
\pagenumbering{gobble}

% A trick to get the bibliography to show up with 1. 2. etc in place
% of [1], [2] etc.:
\makeatletter
\renewcommand\@biblabel[1]{#1.}
\makeatother

% reduce separation between bibliography items if not using natbib:
\let\OLDthebibliography\thebibliography
\renewcommand\thebibliography[1]{
  \OLDthebibliography{#1}
  \setlength{\parskip}{0pt}
  \setlength{\itemsep}{0pt plus 0.3ex}
}

% Set correct font for bibliography (doesn't work yet)
%\renewcommand*{\bibfont}{\bibFont}

% No paragraph indenting to match the VPH format
\setlength{\parindent}{0pt}

% Skip a line after paragraphs
\setlength{\parskip}{0.5\baselineskip}
\onecolumn

% titling definitions
\pretitle{\begin{center}\titleFont}
\posttitle{\par\end{center}\vskip 0em}
\preauthor{ % Fonts are set within \Authors
        \vspace{-1.1cm} % Bring authors up towards title
        \begin{center}
        \begin{tabular}[t]{c}
}
\postauthor{\end{tabular}\par\end{center}}

% Define title, empty date and authors
\title {
%  Competition can account for stopping in a gradient-following model
%  of the retinotectal projection \\
%  \emph{or} \\
%  Competition provides stopping for the chemoaffinity
%  theory which is robust to noise \\
%  \emph{or} \\
%  Self-organisation of topographically ordered axon connections can take place
%  in a noisy environment if axons compete \\
%  \emph{or} \\
Time to stop; agent-based modelling of chemoaffinity with competition
}
\date{} % No date please
\author{\Authors}

% Simplified chemoaffinity explains retinotectal mapping via local axon-axon
% interactions
%
% Focus on the chemoaffinity idea for this model. Show that local interactions
% can provide an account for growth cones finding their targets.

%% END OF PREAMBLE

\begin{document}

\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
\maketitle

\vspace{-1.8cm} % HACK bring the introduction up towards the title. It
                % would be better to do this with titling in \maketitle

%% \color{grey}
%% % Note this is the SfN draft abstract, and will need rewriting
%% % COULD go into the introduction, with some modification.
%% \emph{In the classic Chemoaffinity theory, the retinotectal axon projection is
%% thought to use pairs of orthogonal signalling gradients in the retina to
%% specify the eventual location of synapses made on the surface of the
%% tectum/superior colliculus. Similar orthogonal gradients in the tectum provide
%% a coordinate system which allows the axons to match their prespecified
%% destination with the correct location. Although the Ephrins have been shown to
%% guide axons toward their destination, there has yet to emerge a complete
%% account of the local interactions which halt the axonal growth cones in the
%% correct locations to recreate the topography of the retinal cells. The model
%% of \citet{simpson_simple_2011} provides an account of the basic topographic
%% arrangement of cells on the tectum, as well as reproducing well known surgical
%% and genetic manipulation experiments. However, it suffers from the absence of
%% a local chemotactic guidance mechanism. Instead, each agent in that model is
%% given instantaneous knowledge of the vector that would move it toward its
%% pre-specified destination. In addition to the globally supervised
%% chemoaffinity term, \citet{simpson_simple_2011} introduced a competitive
%% interaction for space between growth cone agents and a receptor-receptor
%% axon-axon interaction in order to account for the full set of experimental
%% manipulations. Here, we propose the replacement of the chemoaffinity term with
%% a gradient following model consisting of axonal growth cone agents which carry
%% receptor molecule expression determined by their soma's location of origin on
%% the retina. Growth cones move on the simulated tectum guided by two pairs of
%% opposing, orthogonal signalling molecules representing the Ephrin ligands. We
%% show that with only the receptor-ligand mediated chemoaffinity term and a
%% receptor-receptor based axon-axon interaction term (meaning that all growth
%% cone interactions are communicated by receptor signalling), a full range of
%% experimental manipulations to the retinotectal system can be
%% reproduced. Furthermore, we show that the observation that competition is not
%% an essential requirement for axons to find their
%% way \citep{gosse_retinotopic_2008} is also accounted for by the model, due to
%% the opposing influences of signalling gradient pairs. Finally, we demonstrate
%% that, assuming exponentially varying receptor expression in the retina, tectal
%% ligand expression should either be exponential if the receptor-ligand signal
%% induces repulsion (i.e. gradient descent) or logarithmic if the signal induces
%% attraction (gradient ascent). Thus, we find that a model analogous to the one
%% we presented in \citet{james_modelling_2020} that accounts for murine barrel
%% patterning is also a candidate mechanism for the arrangement of the more
%% continuous retinotectal system.}
%% \color{black}

\emph{The development of the axonal projection from the retina to the tectum
is an important model for the study of how the brain organises its
connectivity during development. In fish, amphibians and mammals, retinal
ganglion cells grow their axons into the brain and make topographically
ordered connections on the tectum.
%
Computational models seeking to provide a plausible mechanism by which this
system can order itself have existed since the 1970s. However, most of these
have been phenomenological, some only possible with the invokation of a global
supervising process that provides the state of the entire system to each
axonal agent.
%
Here, we investigated whether, using only receptor and ligand interactions, a
purely self-organising model could be found that would successfully reproduce
both the normal development of the projection and its behaviour under the many
experimental manipulations carried out by neurobiologists.
%
We show that a model based on gradient following (via receptor-ligand
interactions) combined with receptor-receptor mediated axonal repulsion, in
which information is only transferred by nearest-neighbour interactions, can
account for a broad range of experimental results.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The retinotectal projection has been an important model system in the study of
how the cells of the central nervous system are accurately connected together
into functional networks. This projection connects the light-gathering cells
in the retina to movement-related cells in the optic tectum (known as the
superior colliculus in mammals). Light sources originating close to each other
in the environment tend to activate retinal ganglion cells (either
directly \cite{iRGC_citation} or indirectly via rods and
cones \cite{rods_cones_citations}) situated close together in the eye so that
an image of the environment is formed on the retinal surface. It has been
discovered that the topography of the retina is preserved within brain regions
that process this information such that cells which are adjacent within the
retina primarily excite cells adjacent in the tectum. Of the retinal axons
that project directly to the tectum, those originating in the temporal retina
make connections in the rostral tectum; thoes originating in dorsal retina
connect to lateral tectum (Fig.\,\ref{}A).  This pattern of connectivity
establishes itself before activity \citep{something}, indicating that during
development there must exist a mechanism which ensures the correct arrangement
of the axons which leave the retina and connect to cells in the tectum.

One reason for the success of the study of the retinotectal projection is its
capacity to be experimentally manipulated. In some non-mammalian species, both
the retina and the tectum can be partially ablated, or even physically
reorganised in-vivo, after which axons regrow to restore the order and
function of the system for the individual animal. Influential work by
R.~W.~Sperry and co-workers during the mid-twentieth century exploited this
manipulability, leading to Sperry's 1963 summary of the \emph{chemoaffinity
theory} \citep{sperry_chemoaffinity_1963} which proposes the existence of
morphogenetic gradients that guide axons to their destination. The
chemoaffinity theory was given robust support by the discovery of the ephrin
ligands and their
receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been
shown to form into graded expression fields in the
retina \citep{braisted_graded_1997},
tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other
sensory systems, such as the somatosensory
system \citep{vanderhaeghen_mapping_2000}. The Ephrin ligands have a clear
effect on axonal outgrowth, as shown
in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004}
and by
in-vivo \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}
studies. \emph{Does this para need finishing?}
%
% A-P:
%
% EphA and ephrinAs: Anteroposterior axis of retiontectal projection
% (Nasal-Temporal on retina; Anteroposterior on tectum).
%
% RGM (tectum) Neogenin (retina)
%
% D-V:
%
% EphB and ephrin B: Dorsoventral on retina; dorsoventral on tectum.
%
% En-2: Repels temporal growth cones of xenopus ret-tec axons and attracts
% nasal ones. Brunet et all, Nature, 2005
%
% Ryk/Fz (retina) and Wnt3 (tectum) (Flanahan, 2006)
%

%It is tempting to consider the retinotectal projection well understood. With a
%comprehensive theory supported by a biochemical mechanism, is there anything
%left to understand? That question can be answered by reviewing retinotectal
%modelling papers.

%Mini-review here, which contrasts some of the modelling papers. The upshot is
%that a central problem with models is not how the axons know how to get closer
%to their destination, but how they know they have \emph{arrived} at their true
%destination. Also make the point that many of the models are phenomenological
%in nature and do not elucidate the mechanisms behind the arrangement.%

In a recent paper, we proposed a self-organising mechanism, based on
morphogenetic signalling gradients, which can arrange
axons growing from the thalamus to the somatosensory cortex into the well
known murine barrel cortex pattern~\citep{james_modelling_2020}. In
characterising that system, we explored the effect of various types of
noise. We found that the mechanism was robust to noise in the expression of
the signalling molecules over a wide range of amplitudes and length-scales,
and that noise in the interaction parameters, which are obtained by sampling
the signalling molecules in the source tissue (the thalamic barreloid field)
could cause topological defects. The question of noise in axon guidance has
been explored by \citet{goodhill_can_2016}.

%% \color{grey}
%% \textbf{New introduction:} Simpson and Goodhill presented a model in which the
%% chemoaffinity mechanism is non-biological. We offer a model in which a
%% competition mechanism counteracts the drive of axons along the ephrin
%% gradient. Will we need the axon-axon interaction mechanism based on
%% non-same-Eph repulsion?

%% \textbf{Another approach:} We present a model of gradient-following based on graded
%% interactions between the growth cones of retinal ganglion cell axons and the
%% tectal surface. We show that a combination of exponentially graded expression
%% in the retina \citep{reber_relative_2004} and exponentially (for repulsively
%% interacting receptor ligand pairs) \emph{or} logarithmically (for attractively
%% interacting receptor/ligands) graded expression on the tectum permits a purely
%% local model to reproduce the majority of reported phenomena associated with
%% retinotectal organisation.

%% There have been many approaches to modelling the organisation of the
%% retinotectal projection. Many of these models are wholly or partially
%% phenomenological.

A model which carefully deals with receptor
binding is given by \citet{naoki_revisiting_2017} \citep[see also][]{mortimer_bayesian_2009}.

Gradient based models: \citet{nakamoto_topographically_1996}

\color{black}
\section{Results}

\subsection*{Chemoaffinity as the sole patterning mechanism}

The first question we tried to answer was whether a gradient following
mechanism, acting alone, is sufficient to pattern the tectum. To do so, we
simulated the retina and tectum as unit square regions and defined retinal
receptor expression and tectal ligand expression across these squares
(Fig.\,\ref{f:ex}B-D). The simulated retina possessed $n$=400 (20$\times$20) retinal
ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square
elements, with defined levels of expression for each ligand. One spatial unit
corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.

\begin{figure}
\includegraphics[width=\linewidth]{./images/expressions_fig.png}
\caption{Simulated retinal and tectal signalling molecule
expression. \textbf{A} A schematic of retino-tectal mapping.
%
\textbf{B} Retinal receptor expression, $r_i$. Four receptors are expressed and
displayed here in two dual-colour maps. $r_0$ and $r_1$ are red and blue;
$r_2$ and $r_3$ are cyan and magenta. $r_0$ increases as an exponential
function of $-x$; $r_1$ increases with an exponential function of $-y$. $r_2$
and $r_3$ have the opposite sense. Retinal receptor expression is modelled
with exponential functions because there is ample evidence to support this
form. $N$: nasal, $T$: temporal, $D$: dorsal, $V$: ventral.
%
\textbf{C} Retinal ligand expression, $l_i$, is expressed in complimentary
gradients \citep{hornberger_modulation_1999}.
%
\textbf{D} Tectal ligand expression $L_i$. Tectal ligand expression is shown
in an exponential form,
although there is less evidence to determine whether tectal expression is
exponential, linear or logarithmic.
$L$: lateral, $M$: medial, $C$: caudal, $R$: rostral.
}
\label{f:ex}
\end{figure}

Each retinal ganglion cell (RGC) projects growth cones (referred to
as \emph{branches}) which carry a set of $N$ receptors, $r_i$ indexed by $i$ and
expressed at levels determined by the cell soma's location on the
retinal grid. ($N$ ligands, $l_i$, are also expressed by RGCs, but these are not
used by the chemoaffinity simulation.) The expression of each receptor
varies only with respect to one spatial dimension.

The minimum number of gradients required to provide a coordinate system on the
surface of the tectum within which axons could find a target location is
$N=2$, arranged in orthogonal directions. However if there are only two
gradients, imagining that one guides axons laterally and the other guides
axons in the caudal direction, then it is necessary for there to be an
additional `stopping mechanism' to prevent all of the axons eventually being
guided to the most lateral/caudal position. To avoid the need for a different
stopping mechanism, we chose to model \emph{opposing pairs of orthogonal
gradients} with $N=4$. With opposing gradients, an axon which interacts most
strongly with a gradient in the caudal direction would be drawn to the caudal
side of the tectum, whereas an axon whose strongest interaction was to climb a
gradient in the rostral direction would be drawn to the rostral side. There is
evidence that multiple gene expression gradients are found in the brain, with
examples of opposing pairs including XX and YY.

We defined 4 receptor expression gradients arranged in orthogonal pairs
with the gradient of $r_0$ being orthogonal to that of $r_1$. $r_2$, whose
gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:expressions}B).

Because there is convincing evidence that EphA and EphB receptors are
expressed in exponentially increasing
patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
we use an exponential form for retinal receptor expressions, in common with
other modelling
studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
We adopt the same precise form for the retinal receptor expression
as \citet{simpson_simple_2011} which is
\begin{equation} \label{e:retrcpt}
r_i(x,y) = 1.05 + 0.26 \exp(2.3 u)
\end{equation}
Where $u$ gives the direction with which the expression increases and is
substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for
$r_3$ (Fig.\,\ref{f:ex}B).

Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in
orthogonal pairs of gradients. Although several studies model tectal ligand
expression with exponential functions \citep{koulakov_stochastic_2004}, the
experimental evidence for the form of tectal ligand expression is more ambiguous than that for
retinal receptor expression (Fig.\,\ref{f:expressions} summarises results from
the literature). We kept in mind the
possibility that tectal ligand expression may be modelled by exponential,
linear or logarithmic functions, but initially, we set it to the same
exponential function used for retinal expression,
\begin{equation} \label{e:tecligexp}
L_i(x,y) = 1.05 + 0.26 \exp(2.3 u),
\end{equation}
and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain
the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$
(Fig.\,\ref{f:ex}D).

While we do not explicitly name $r_0$, $r_1$,
etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes EphA,
EphB, Ryk \citep{schmitt_wntryk_2006} and
Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ includes the
ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and
RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a
r\^ole in retinotectal map formation. The fact that we have chosen opposing
gradients is also guided by experimental observations that these exist in the
brain \citep{opposing_gradients}. Without a system of opposing gradients, it
would not be possible for axons to stop without an additional signal
instructing them to do so.

%already said that:
%Figure~\ref{f:ex} shows colour maps of receptor and ligand
%expression. Coordinates within the simulated retina and tectum are normalised,
%with 1 corresponding to 2 mm on the tectum and on the retina for
%mouse \citep{reber_relative_2004}.

%% This means that growth cones of approximate
%% radius 5 $\mu$m = 0.005 mm \citep{goodhill_can_2016} have a radius $r$ =
%% 0.0025 in the arbitrary units of the simulation. For a realistic simulation
%% based on mouse, the approximate RGC density in the retina is around 44000
%% cells per retina \citep{jeon_major_1998}. This sets the side length of the
%% retina to $\sqrt{44000} \approx 200$. This results in a simulation of
%% prodigious computational size, which although tractable (10 s per step,
%% requiring several hours, rather than days to run) is not convenient. To
%% maintain the relationship between growth cone size and cell density, we permit
%% a smaller number of growth cones ($400\times8=3200$) but keep the fractional
%% area of the branches at 0.86 ($44000 \times \pi r^2$) (if there is one cone
%% per axon) or 6.9 (if there are 8 cones per axon).

Each discrete, square element in Fig.\,\ref{f:ex}B projects one RGC axon with $M=4$
branches to the tectum, carrying the receptor expression at the retinal
element's location.
At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, at time
$t$ of each branch $b$ is updated according to
%
\begin{equation}
\mathbf{x}_{b,t+1} = \mathbf{x}_{b,t} + \mathbf{M}_{b}
\end{equation}
%
where $\mathbf{M}_{b}$, the branch movement vector, is the sum of a chemoaffinity
effect and a `border effect', $\mathbf{B}_b$, which acts to retain branches
within the tectal region:
%
\begin{equation} \label{e:mv}
\mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b + 0.5 \mathbf{B}_b
\end{equation}
%
evaluated at time $t$.

% G - chemoaffinity description
%\subsection{Chemoaffinity}
The chemoaffinity component, $\mathbf{G}_b$, is governed by receptor-ligand
signalling. The signal transmitted when a tectal ligand binds to a RGC
receptor on a branch (forward signalling) can lead to one of two effects; the
signal may induce an attraction towards the ligand-expressing region or a
repulsion away from it.
%
For axon-tectum interactions, we implemented attraction as a climbing of the
tectal ligand expression gradient and repulsion as gradient descent.
%
To determine the strength of the effect we assumed a purely linear receptor
binding model, and set the chemotactic movement vector of the branch $b$ at
location $\mathbf{x}_b$ on the tectum to be
%
\begin{equation}\label{e:G}
\mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
\end{equation}
%
where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor
pair $i$, $\nabla L_i$ is the gradient of expression of ligand $i$ on the
tectum and $F_i$ denotes the direction of the interaction induced when a
molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes the value
$-1$ for a repulsive interaction or $1$ for an attractive interaction.
%
We assumed that all
receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as
for EphA-ephrinA
coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
%
$m_{\!_G}$ in Eq.\,\ref{e:mv} is a
scalar parameter which controls how much movement is generated for a given
chemotactic movement vector, $\mathbf{G}_b$.
% end chemoaffinity description

% \subsection{Border effect} % border effect description
The border effect is also based on gradient following by assuming that
there is some other molecular signal which acts on all branches near the
boundary of the tectal tissue. For a branch with position $(x,y)$, $\mathbf{B}_b$ is
given by:
%
\begin{equation}
B_{b,x} = \begin{cases}
        r-x      & x<r \\
        1-r-x    & x>1-r
\end{cases}
B_{b,y} = \begin{cases}
        r-y      & y<r \\
        1-r-y    & y>1-r
\end{cases}
\end{equation}
%
Thus $\mathbf{B}_b$ is equivalent to the action of a repulsive signalling
molecule expressed around the border of the tectum, whose expression increases
quadratically outside the tectum and affects any branch touching (or outside)
the boundary.
% end border effect description

% \subsection*{Initial conditions}
Branches were initially randomly distributed in a stripe at the rostral side
of the tectum at the start of each simulation. Each RGC axon $k$ was assigned a
random initial position coordinate:
\begin{equation}\label{e:ic}
\mathbf{x}_{k,t=0} = (U(0,1), U(-0.2,0))
\end{equation}
where U(p,q) is a number selected from a random uniform distribution in the
range $[p,q)$. Each of the 4 branches per RGC axon was given its parent
axon's initial position, plus a randomly generated offset with coordinates
derived from a normal distribution of mean 0 and standard deviation 0.1.
% end initial conditions

The results of running this simulation for 300 time steps are shown in
Fig.\,\ref{f:G}. Fig.\,\ref{f:G}A indicates the colour code used for retinal
cell position; dorsal retinal cells are indicated by green, nasal cells by
red. Lightness gives position along the relevant axis so that cells at the
dorsal-nasal corner are yellow and those at the ventral-temporal corner are
black. Fig.\,\ref{f:G}E shows the expected layout of retinal ganglion cell
axons on the tectum. The axons retain the colour of their originating cell on
the retina. Nasal cell axons should be arranged
caudally; dorsal axons laterally; axons should be evenly spaced and the
pattern should span the entire tectum.  Fig.\,\ref{f:G}, panels B--D show
`fish net' plots of the average location of the 4 branches of each axon (the
axon's centroid) at three timepoints. Fig.\,\ref{f:G}B indicates the
randomised placement of axon branches at the rostral end of the tectum as
specified by the initial conditions (Eq.\,\ref{e:ic}). In these plots, lines are drawn between
axons that are expected to arrange themselves adjacently to one another by the
end of the simulation. The number of line crossings in this plot gives an
indication of the disorder in the arrangement. Fig.\,\ref{f:G}G shows the positions of the individual branches
(i.e.~growth cones) at $t=300$ that contribute to the plot in
Fig.\,\ref{f:G}D. By this time, the topological order matches that shown in
Fig.\,\ref{f:G}E. Opposing gradients have sorted the axons along the two axes
to form a grid of cells. However, the arrangement does not span the full
tectum. Fig.\,\ref{f:G}F shows two pattern metrics; $\epsilon$ is the mean
distance error of the axon centroids, given by:
%
\begin{equation}\label{e:eps}
\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
\end{equation}
%
where $\mathbf{x}_{k}$ is the location of the centroid of the
branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon
$k$. $n$ is the number of RGC axons in the system.
%
$\epsilon$ tends to 0 as the pattern approaches the target layout. However,
two patterns may have the same non-zero value for $\epsilon$, despite one
being qualitatively more disordered than the other. For this reason, we also
use the number of crossings in the fish net plot of Fig.\,\ref{f:G}D as a
second metric, called $\eta$, which tends towards an expected target number of
crossings, $\eta_t$ as the pattern approaches the target pattern.  For the wildtype
pattern, $\eta_t=0$, as in Fig.\,\ref{f:G}E, although note that in some patterns,
$\eta_t$ is non-zero. Fig.\,\ref{f:G}H shows selected
axons (one from the centre and one from each corner of the retina) and their
position history.

% ./build/sim/agent/agent1 configs/simpler/m_ee_G.json configs/simpler/e_wt_fig2.json
\begin{figure}
\includegraphics[width=\linewidth]{./images/j4_ee_G_wt_fig2_exit_true.png}
\caption{The pure gradient-following model. \textbf{A} The origin position
on the tectum gives axons a colour according to a dual-colour map. Axons in
panels B-H that are red are those which originated in the most naso-ventral
corner of the square simulated retina. \textbf{B}-\textbf{D} Axon centroid
positions on the simulated tectum (dashed line) at time 0, 40 and 300
units. Lines show adjacency relationships between the axons' origin positions
on the retina. \textbf{E} Layout of axon centroids on the tectum observed in
experimental results. \textbf{F} The sum-of-squared difference error,
$\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon
arrangement plotted with respect to simulation time. \textbf{G} The
arrangement of axon branches at t=300. \textbf{H} Selected axon locations at
t=300, with path histories in grey.  \emph{Abbreviations: D dorsal, V ventral, T
temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
\label{f:G}
\end{figure}

The form of the retinal receptor expression (Eq.\,\ref{e:retrcpt}) is based on
empirical measurement, but the tectal ligand expression
(Eq.\,\ref{e:tecligexp}) was chosen arbitrarily. To expand the pattern in
Fig.\,\ref{f:G}D so that it covers the entire tectum, we needed only to
adjust the tectal ligand expression, substituting the alternative exponential
expression:
%
\begin{equation} \label{e:tecligexp2}
L_i(x,y) = 1.05 + 0.26 \exp(1.1 u)
\end{equation}
%
for Eq.\,\ref{e:tecligexp}.

The result at $t=300$ is shown in Figs.\,\ref{f:G0}A
and \ref{f:G0}B. The axon centroids are arranged across the full tectal
surface.  The arrangement is slightly disordered around the middle of the
tectum. This is caused by numerical artefacts from the discretization of the
ligand expression and the fact that in the centre, the difference between the
opposing receptor-ligand signals is very small. Nevertheless, the result
demonstrates that a modified form for the tectal ligand expression (which was
found easily by hand tuning) allows a well ordered map to form for the
wildtype system. Fig.\,\ref{f:G0}C shows the error metric, $\epsilon$,
which is much closer to 0 with the new ligand expression.

% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_wt_fig3.json
\begin{figure}
\includegraphics[width=\linewidth]{./images/j4_eE_G_wt_fig3_exit_true.png}
\caption{By tuning the tectal ligand expression (replacing Eq.\,\ref{e:tecligexp} with
 Eq.\,\ref{e:tecligexp2}), the pure gradient-following model can achieve a
 successful layout in which the axons are arranged evenly across the full
 tectum. At the centre of the tectum, where neither retinal receptor
 expression nor tectal ligand expression is high, the ordering is weaker than
 towards the edge. \textbf{A} Axon branches \textbf{B} axon
 centroids \textbf{C} The pattern metric $\eta$ is zero, as it was in
 Fig.\,\ref{f:G}F and $\epsilon$ approaches a much lower value as $t \to \infty.}
\label{f:G0}
\end{figure}

This `chemoaffinity only' model begins to encounter problems when we introduce
simulations of experimental surgical manipulations. Simulated surgery is
performed by manipulating the maps shown in
Fig.\,\ref{f:ex}. Fig.\,\ref{f:trot90} shows the expression maps and the
resulting gradients that result from the surgical manipulation in which a
graft of the tectum is rotated by a quarter turn and replaced. We rotate an
8$\times$8 square of the tectum and then recompute the gradients. As the ligand
expression is discretized rather than continuous, we compute the gradient
numerically, with the gradient in direction $x$ computed as the difference
between the expression of the two neighbouring elements divided by twice the
element distance. As there are four ligands expressed on the tectum, there are
four pairs of gradient maps, each of which is shown in Fig.\,\ref{f:trot90}.

% Visualisation of ligand expression and gradients on the tectum for the
% graft-rotated 90 degrees simulated surgical manipulation
% ./build/sim/agent/tissuevis configs/simpler/m_eE_G.json configs/simpler/e_tecrot90_fig4row1.json
%
\begin{figure}
\includegraphics[width=\linewidth]{./images/Tissuevisb.png}
\caption{Ligand expression and gradients on the tectum for the
graft-rotate 90 degrees simulated surgical manipulation}
\label{f:trot90}
\end{figure}

Fig.\,\ref{f:tmanip} shows the tectal ligand expression for ligands $L_1$ and
$L_2$ for the five other surgical manipulations applied in this work. These
are a 180\degree~rotation (Fig\,\ref{f:tmanip}A), a graft-swap manipulation
(Fig\,\ref{f:tmanip}B), ablation of half the retina (Fig\,\ref{f:tmanip}C),
and ablation of half the tectum (Fig\,\ref{f:tmanip}D). A mismatch
manipulation was also performed in which which half of the retina is removed
along with the half of the tectum to which the remaining retina would project.

Fig\,\ref{f:Gsurg} shows the results of 6 simulated surgical
manipulations. In A, a patch of tectal tissue has been cut out and rotated by
a quarter turn (the manipulation illustrated in Fig.\,\ref{f:trot90}). For
this simulation, the figure shows i) The locations of the axon centroids at
simulation end, ii) (inset) the expected final layout of the tectum, according
to the results of experimental work, with the colour code giving the origin
location of the axon's soma on the retina and iii) the error metric,
$\epsilon$ plotted versus simulation time. Simulation time was 1000 time
steps. The final locations of branches and axon centroids indicates that a
qualitative rotation of the graft patch has occurred, with red branches inside
the patch being located at the lateral-caudal corner rather than the
medial-caudal corner. A portion of the graft region (the medial side, after
rotation) is free of branches, indicating that the pattern is partial. A
similar result is shown for the 180\degree~example in B, in which the caudal
half of the rotated patch is free of branches, but the rostral half contains
branches with the correctly rotated orientations.

Fig.\,\ref{f:Gsurg}C shows the result of the patch swap manipulation, in
which 12$\times$4 element grafts, one from the caudal side of the tectum and one from
the rostral side are swapped. This simulation is unsuccessful in patterning
the caudal half of the tectum. The large gradients, both positive and
negative, caused by the discontinuities along the long edges of the grafts
attracts almost all the branches (whose initial position is the rostral edge)
and very few pass through to the caudal side.

The retinal ablation shown in Fig.\,\ref{f:Gsurg}D is also unsuccessful, with
the axons from the surviving nasal retina (the temporal half was removed)
failing to arrange themselves across the entire tectum, due to the pure
`position specification' which the dual gradient model achieves. Here, the
average axon position on the rostral-caudal axis of the tectum is plotted versus the average
nasal-temporal origin of the axon on the retina. Similarly,
the tectal ablation simulation fails to reproduce experimentally observed
results in which the retinal map `squashes up' within the remaining tectal
region. Instead, axons from the temporal retina arrange with normal spacing
and axons from the nasal retina are pushed to the caudal-most side of the
remaining tectum (Fig.\,\ref{f:Gsurg}E). A similar result occurs in the
mismatch experiment (Fig.\,\ref{f:Gsurg}F), but in this case there are no
axons from the temporal retina, which was ablated and only the axons
from the nasal retina are present, arranged in a line.

% The chemo manipulations figure is made up from 4 figures generated with
% graph_layout::g
% (see scripts/fig_G_surgical.sh)
\begin{figure}
\includegraphics[width=0.95\linewidth]{./images/fig_chemo_manipulations.png}
\caption{Surgical manipulations in the chemoaffinity-only model. \textbf{A} 90 degree
rotation. \textbf{B} 180 degree rotation. \textbf{C} Graft swap. \textbf{D}
Retinal ablation \textbf{E} Tectal ablation \textbf{F} mismatch ablation}
\label{f:Gsurg}
\end{figure}

Together, these results confirm that, while a pure gradient-following
implementation of Sperry's chemoaffinity theory is able to reproduce the
wildtype tectal arrangement of retinal ganglion cell axons, it cannot account
for the results of surgical manipulation experiments.

\subsection*{Combining chemoaffinity with competition}

Previous work has shown that competition can contribute to the correct layout
of the retinotectal system \citep{stuff} and to the patterning of other
systems including the murine cortical whisker
barrels \citep{james_modelling_2020} \textbf{(FIXME owt else?)}.

\citet{simpson_simple_2011} showed that repulsive interactions combined with a
chemoaffinity-like effect could account for the surgical manipulations shown
in Figs.\,\ref{f:trot90}--\ref{f:Gsurg}. To confirm that a competitive
mechanism could operate successfully in combination with the
gradient-following chemoaffinity mechanism described above, we first
implemented the `space based' competition of \citet{simpson_simple_2011},
updating the movement vector with a term $m_c \mathbf{C}_b$:
\begin{equation} \label{e:mv2}
 \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b +  m_c \mathbf{C}_b + 0.5 \mathbf{B}_b
 %%+  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
\end{equation}
where $m_c$ is simply a scalar parameter.
%
%
%% introduced
%% three competition mechanisms and investigated their behaviour.
%
%% These were: space-based competition (`$\mathbf{C}$'), as described
%% in \citet{simpson_simple_2011}; a receptor-receptor signalling mechanism
%% (`$\mathbf{I}$'), similar to that described in \citet{simpson_simple_2011},
%% and a receptor-ligand mass-action signalling model
%% (`$\mathbf{J}$'). The movement vector (Eq.\,\ref{e:mv}) becomes
%% %
%% \begin{equation} \label{e:mv2}
%% \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b +  m_c \mathbf{C}_b +  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
%% \end{equation}
%% %
%
%% Although two of these competitive mechanisms were investigated
%% by \citet{simpson_simple_2011}, the first (space-based competition) offers no
%% mechanism by which axon growth cones are able to communicate and the second
%% mechanism was implemented only for a single axis in the model (EphA receptors,
%% expressed in a gradient along the XX retinal axis). We wondered whether a
%% \emph{single} signalling mechanism working alongside the `untuned'
%% chemoaffinity with ligand expression given by Eq.\,\ref{e:tecligexp} could
%% successfully reproduce the wildtype and manipulation results. Alongside the
%% same chemoaffinity mechanism, we studied each competitive mechanism in turn,
%% setting $m_{\!_I}=m_{\!_J}=0$ to give a `$\mathbf{GC}$' model,
%% $m_c=m_{\!_J}=0$ to give a `$\mathbf{GI}$' model or $m_c=m_{\!_I}=0$ to give a
%% `$\mathbf{GJ}$' model.
%
%%\subsection*{Chemoaffinity plus space-based competition (GC)}
%
The competition movement vector, $\mathbf{C}_b$, introduces a repulsive effect
acting between any two branches, whether they originated from different
retinal ganglion cells or from the same cell. An assumption is that there is
some kind of self-other signal which acts to discourage two cells from
approaching each other. $\mathbf{C}_b$ is defined as a proximity-weighted sum
of the repulsion due to all nearby branches:
%
\begin{equation} \label{e:C}
\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb} W_c
\end{equation}
where $\hat{\mathbf{x}}_{kb}$ is the unit vector from branch $k$ to branch
$b$. The distance-dependent weight $W_c$ is given by
\begin{equation}
W_c(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r_c}   & d_{kb} \leq 2r_c \\
     0 & d_{kb} > 2r_c
     \end{cases}
\end{equation}
for two growth cones of interaction radius $r_c$ separated by a distance
$d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
branch $b$ and the total number of branches in the system is the number of RGC
axons, $n$ multiplied by the number of branches per axon, $M$. This
description of the competition matches that
in \citet{simpson_simple_2011}. Because the model incorporates a chemoaffinity
effect and a competition effect, we called it the `$\mathbf{GC}$ model'.

The model defined by the receptor and ligand expressions of Eqs.\,\ref{e:retrcpt}
\& \ref{e:tecligexp} along with Eq.\,\ref{e:mv2} has three
free parameters; $m_{\!_G}$, $m_c$ and $r_c$. To find the best combination of
these parameters, we reasoned that evolution would optimize the arrangement of
the wildtype pattern. We performed a simulated annealing optimization to
minimize the error metric, $\epsilon$, for the wildtype pattern
(Fig.\,\ref{f:G}E) at simulation time 1000. The parameters returned were
$m_{\!_G}$ = 0.003841, $m_c$ = 0.09959 and $r_c$ = 0.39918. Fig.\,\ref{f:GC}
shows the $\mathbf{GC}$ simulation with these parameters. Fish net plots at
$t=40$ and $t=1000$ show that the pattern develops quickly and compares well
with the expected layout (Fig.\,\ref{f:G}E). We investigated the response of
this model to the same set of surgical manipulations shown for the
chemoaffinity-only model in Fig.\,\ref{f:Gsurg}. We confirmed that the
addition of competition leads to a better correspondence to the experimental
results than for the chemoaffinity-only model (see supplementary
Fig.\,\ref{f:GCsurg}).

\subsection*{Axon-axon interactions}

Because the competition effect, $\mathbf{C}_b$ (Eq.\,\ref{e:C}), does not
describe a signalling mechanism, and because \citet{simpson_simple_2011}
reported that an additional mechanism is required to explain the gene
manipulation experiments of \citet{brown_topographic_2000}
and \citet{reber_relative_2004}, we looked for a receptor-ligand interaction
that could signal between axon branches and enable them to adjust their
movements away from one another.

% First, look at I, from Simpson and Goodhill.
First, we took inspiration from Simpson and Goodhill's model for the
interaction between EphA receptors, which is defined (with receptor $r_i$
representing EphA) as:

%We modelled interactions between axon growth cones due to signalling between
%receptors of the same type. For these interactions, repulsion caused a
%movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
%branch $k$ to branch $b$ causing them to move further apart; attraction caused
%the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
%is given by
%
\begin{equation}
\mathbf{I}_b
= \frac{1}{|B_{b,I}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
/ r_{i,k} > 1.1
\end{equation}
%
$r_{i,k}$ is the expression of receptor $i$ on branch $k$ and the weight here
is defined as a function of the distance, $d_{kb}$, between branches:
%
\begin{equation} \label{e:WI}
W_{\!_I}(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r_{\!_I}}   & d_{kb} \leq 2r_{\!_I} \\
     0 & d_{kb} > 2r_{\!_I}
     \end{cases}
\end{equation}
%
$B_{b,I}$ is the set of growth cones within a distance $2r_{\!_I}$ of branch
$b$. $\hat{\mathbf{x}}_{kb}$ is as defined in Eq.\,\ref{e:C}. This mechanism
adds a weighted repulsion if the ratio of receptor expression on
branch $b$ and the expression on branch $k$ exceeds the threshold of 1.1.

Simpson and Goodhill introduced this interaction as an additive modification
to the main, distance based competition mechanism of Eq.\,\ref{e:C} (adding a
term $m_{\!_I} \mathbf{I}_b$ to the equation). They introduced an interaction
varying along the temporal-nasal retinal axis and mediated by one receptor,
EphA, for which there was experimental data supporting a repulsive
effect. They noted that the use of a receptor ratio (comparing the ratio with
a `discrimination limit', above which repulsion would be activated) is
equivalent to a multiplicative interaction between EphA receptors and ephrin-A
ligands with the strong assumption that ligand expression is the inverse of
receptor expression, i.e. that $l_i = 1/r_i$.

We asked i) whether this mechanism could \emph{replace} the non-biological
competition, $\mathbf{C}_b$, and ii) if cast in terms of receptor-ligand
interactions, could the strong ligand expression assumption be relaxed?

% (supp?) Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.

% Next, replicate with receptor-ligand interaction.

% OR

% Look at receptor-ligand interaction, and just point out that the

In order to replace the distance-based competition, $\mathbf{C}_b$, we first
point out that many of the ligands, $l_i$, expressed on the tectum are also
expressed by retinal ganglion cells and so where there is signalling between
tectal and RG cells, there is also the possibility for inter-RGC
signalling. Furthermore, if a receptor-ligand interaction on the tectum leads
to growth cones moving away from a region of high ligand expression, then the
equivalent axon-axon interaction would also be expected to be repulsive.

Taken alone, the EphA forward interaction between one RGC branch, $b$, and
another, $k$, will cause a repulsion if $k$ originated more nasally than
$b$. A branch originating on the \emph{temporal} side of $b$ will cause no
repulsion of $b$. However, although $k$ does not induce repulsion in $b$, $b$
should induce a repulsion in $k$ and thus, it could be argued that only a
single ephrin-A ligand gradient should be required in the retina to ensure
that repulsive signals are transmitted between axons whose cell bodies
originate far apart in the retina. However, there is an asymmetry in the
interaction. Consider a cluster of growth cones, $k$, originating from the
nasal retina near to a single cone, $b$, from the temporal retina and assume
that $r_b/r_k>1.1$. The single cone is affected by the sum of 3 interactions
and has a movement of 3 units added to its movement vector (in the temporal
direction). Now reverse the originating locations so that the cluster, $k$,
originates temporally and the individual cone, $b$, comes from the nasal
retina. Now $r_k/r_b>1.1$ and each member of the cluster experiences an
addition of just one unit to its movement vector and the cones move apart at a
speed that is only $1/3$ of that for the original origins.

To avoid this asymmetry, we allowed two receptor-ligand pairs, which are
expressed in opposing directions, to contribute to axon-axon interactions,
ensuring that interactions between single growth cones and clusters are not
dependent on their origin locations. Fig.\,\ref{f:ex}C shows ligand
expressions representing an ephrin-A like expression ($l_0$) and an opposing
expression ($l_2$). We adopted a form for the ligand expression such that
$r_i \times l_i = 4.7727$, with the scalar constant chosen so that the retinal
ligand expression, $l_i(x,y)$ has approximately the same start and end values
as the exponential retinal receptor expression:
%
\begin{equation}
 \label{e:retligexp}
l_i(x,y) = \frac{4.7727}{1.05 + 0.26 \exp (2.3 u)}
\end{equation}
$u$ was substituted with $(1-x)$ for $l_0$ or $x$ for $l_2$ (Fig.\,\ref{f:ex}C).
%
%With a suitable choice of threshold, $s$, branches originating from the same
%location on the temporal-nasal axis will not experience competitive
%interactions.

Note that it would be possible to postulate two additional retinal ligand
expression fields (for $l_1$ and $l_3$), but we found this not to be
necessary; a single opposing pair of fields can ensure that most pairs of
axons are able to signal the necessity for a competitive interaction. We set
$l_1 = l_3 = 0$.

% Now describe J model, and show the result.
We defined a movement vector arising from a multiplicative receptor-ligand
interaction, $\mathbf{J}_b$ (scaled by a parameter $m_{\!_J}$) to define the
movement vector, $\mathbf{M}_b$, for the `$\mathbf{GJ}$ model' as:
%
\begin{equation} \label{e:mv3}
 \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b + m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
\end{equation}
%
% noting that $m_J J_b$ \emph{replaces} $m_c C_b$.
%
$\mathbf{J}_b$ is defined as
%
\begin{equation}
\mathbf{J}_b = \frac{1}{|B_{b,J}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W_{\!_J}\,Q_{\!_J}(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k})
\end{equation}
%
where $W_{\!_J}$ is distance based weighting analogous to $W_{\!_I}$ and
$B_{b,J}$ is analogous to $B_{b,I}$. $Q_{\!_J}$ determines if the signals
between a branch $b$ expressing receptors $\mathbf{r}_b$ and branch $k$
expressing ligands $\mathbf{l}_k$, separated by a distance $d_{kb}$ will cause
$b$ to move away from $k$. $Q_{\!_J}(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k})$
can be expressed as
%
\begin{equation}
Q_{\!_J}(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}) = \begin{cases}
                 0 & \mathrm{if}~-F_i\,r_{i,b}\,l_{i,k} <
                 s,\,\forall{i}~\mathrm{or}~d_{kb} > 2r_{\!_J} \\
                 1 & \mathrm{otherwise}
     \end{cases}
\end{equation}

where $r_{\!_J}$ is the receptor-ligand interaction radius, $l_{i,k}$ (the
expression of ligand $i$ on branch $k$) is the $i^{\mathrm{th}}$ element of
$\mathbf{l}_k$ and $r_{i,b}$ is the $i^{\mathrm{th}}$ element of
$\mathbf{r}_b$. $F_i$, defined in Eq.\,\ref{e:G}, is -1 if the $r_{i}\,l_{i}$
interaction signals repulsion and +1 if it signals attraction. $s$ is a
positive signal threshold whose value should be around 4.7727; if higher, then
branches whose retinal origin is close will not experience a competitive
interaction. If $s\leq 4.7727$, then \emph{any} pair of branches will compete
as they approach each other. With the constraint $s>0$ comes the assumption
that attractive interactions will have no effect on the competitive movement
vector---a receptor-ligand pair that signals attraction cannot cause
$Q_{\!_J}$ to take the value 1. $Q_{\!_J}$ is set to 1 only if at least one
repulsive signal exceeds the threshold, $s$ and the distance from branch $b$
to branch $k$ is smaller than $2 r_{\!_J}$.

Note that $\mathbf{J}_b$ would be equivalent to $\mathbf{I}_b$ (with $s=1.1$)
if there was a single $l_i$ expressed in the retina and its form was $1/r_i$.

% Now the simulation.
Adopting retinal expressions given in Eqs.\,\ref{e:retrcpt}
and \ref{e:retligexp}, we performed a simulated annealing optimization to find
the best parameters for the $\mathbf{GJ}$ model.  With $m_{\!_G}$ fixed at the
value 0.0012, we allowed the threshold parameter $s$ to vary along with
$m_{\!_J}$ and $r_{\!_J}$. The parameters best able to recover the wildtype
pattern were $s=4.732$, $m_{\!_J} = 0.247$ and $r_{\!_J} = 0.079$, with the
wildtype result shown in Fig.\,\ref{f:GJ}. Note that the discovered value for
$s$ is \emph{less} than 4.7727 meaning that for any pair of branches, either
$r_0\times l_0$ or $r_2\times l_2$ will exceed $s$ and branches all
compete. With $s < 4.7727$ the $\mathbf{GJ}$ model is equivalent to the
$\mathbf{GC}$ model. However, the parameters found for the $\mathbf{GJ}$ model
did lead to better results for surgical manipulations, as shown in
Fig.\,\ref{f:GJsurg}. Values for the pattern metric, $\epsilon$ are
consistently lower than those for the $\mathbf{GC}$ model shown in
Fig.\,\ref{f:GCsurg}. Given that the \emph{systems} are equivalent, the
differences must have arisen due to the optimization process, which must have
found alternative `good' parameters. We verified this by copying the
parameters found for the $\mathbf{GJ}$ model into the $\mathbf{GC}$ model,
setting $m_G = 0.0012$, $r_c = 0.079$ and $m_c =
0.247$. Fig.\,\ref{f:GClikeGJsurg} shows the result for surgical
manipulations. Their equivalence to Figs.~\ref{f:GCsurg} is evident by
inspection. Nevertheless, we retain the receptor-ligand signalling of the
$\mathbf{GJ}$ model as this maintains the ability of the model to respond to
the genetic manipulations described in the next section.

% ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp3.json
\begin{figure}
\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_figcomp3_exit_true_steps_1500.png}
\caption{Chemoaffinity and receptor-ligand mediated competitive interactions
($\mathbf{GJ}$ model). }
\label{f:GJ}
\end{figure}

% ./scripts/fig_GJ_surgical.sh
\begin{figure}
\includegraphics[width=\linewidth]{./images/fig_GJ_surgical.png}
\caption{$\mathbf{GJ}$ model, surgical manipulations}
\label{f:GJsurg}
\end{figure}

\subsection*{Knockin and Knockdown manipulations}

In the early 2000s, a series of genetic experiments added to the set of
manipulations that a model of the retinotectal system would need to
explain. First, \citet{brown_topographic_2000} showed that a selective RGC
knock-in of EphA3 receptors led to an interesting modulation of the maps found
in the tectum. For the strongest knock-in (a homozygote manipulation) two maps
were formed, with knock-in cells finding termination zones more rostral than
their wildtype retinal neighbours. By enhancing the quantity of EphA receptors
expressed by half of all retino-tectal axons, they demonstrated the importance
of retinal receptor expression for the arrangement of the tectal map. In the
weaker, heterozygote manipulation, the maps were still separated, but the
separation `collapsed' towards the rostral end of the tectum
(see \citet{brown_topographic_2000}, Fig\,5). This led to a second study
showing that with an additional knock-down of EphA4 receptors in \emph{all}
RGCs, the retinal origin of the collapse point was shifted temporally and all
heterozygote EphA3 knock-in cells were pushed even further rostrally
(\citet{reber_relative_2004}, their Fig.\,3e). Does the $\mathbf{GJ}$ model
reproduce the results of Brown et al and Reber et al?

\begin{figure}
\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
\caption{$\mathbf{GJ}$ model, Brown et al./Reber et al.~manipulations.}
\label{f:GJeph}
\end{figure}

An assumption made in these papers is that all EphA receptor subtypes
contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which
increases from a low level at the nasal side of the retina, to a high level at
the temporal side. The precise contribution of each subtype to the overall
level is assumed not to be important. In our model, receptor $r_0$ increases
in the retina in a nasal to temporal direction (Fig\,\ref{f:expressions}B) and
so we will manipulate $r_0$ to recreate the Brown et al.~and Reber et
al.~experiments. In half of the RGCs, we knock-in EphA3 by adding to the value
of $r_0$ in these cells either 0.6 (heterzygote EphA3 ki/+) or 3.0 (homozygote
EphA3 ki/ki). To add the additional knock-down of EphA4
in \citet{reber_relative_2004}, we subtract 0.35 from $r_0$ in \emph{all}
RGCs. Fig.\,\ref{f:GJeph}A-C shows, for the single EphA3 knock-in (EphA3
ki/+), branch positions, axon centroids and final rostral-caudal position on
the tectum plotted versus the originating nasal-temporal retinal
coordinate. In A, branches are coloured in the red/green duochrome scheme
according to their origin location on the retina. In B \& C, axon centroids of
RGCs which were treated with the knock-in manipulation are coloured red,
wildtype cells are blue. The map of treated cells on the tectum is shifted
slightly rostrally, but the shift becomes statistically insignificant at about
0.75 along the nasal-temporal axis. While the map collapse is less abrupt than
in the experimental result of \citet{brown_topographic_2000}, it is comparable
with the result from \citet{simpson_simple_2011}.  Fig.\,\ref{f:GJeph}D-F
shows the same trio of graphs for the homozygote double knock-in, EphA3
ki/ki. Here, the treated axons are pushed further towards the rostral end of
the tectum and the untreated axons are pushed caudally. The maps are
completely separated; as in the experiment, there is no `collapse point'.

The rostral movement of treated axons occurs due to their increased expression
of EphA ($r_0$), which increases the interaction with the tectal
rostral-caudal ligand gradient. The opposing receptor expression ($r_2$) is
unchanged by the manipulation and so those axons with increased $r_0$
expression are pulled to a more rostral position than they would otherwise
take. The caudal movement of wildtype axons, whose receptor expression in
Figs.\,\ref{f:GJeph}A-F is unchanged, occurs due to competition with the
caudally-drawn treated axons.
%
%Interestingly, the feature of a slight 'elbow' in the double knockin
%graph (F) at about 0.7 on the nasal-temporal axis matches that seen in the
%experimental result. Well, kinda.

Figs.\,\ref{f:GJeph}G-I show the results for the Reber et al.~manipulation, in
which the EphA expression is selectively knocked in in 50\% of cells and
knocked down in all cells. The collapse point has moved from about 0.75 in
Fig.\,\ref{f:GJeph}C to about 0.85 (\textbf{check}), in agreement with the
experimental result. However, the significant rostral movement of the treated
axons seen in \citet{reber_relative_2004}, where not a single treated axon was
found more than 30\% along the rostral-caudal axis, was absent from our result.

%Considering just one receptor-ligand pair, for instance $r_0$/$l_0$, and
%defining their interaction as repulsive, an axon from the transverse side of
%the retina, which expresses a high level of receptor $r_0$ will interact most
%strongly with an axon from the nasal side of the retina, where $l_0$ is most
%highly expressed. The interaction with axons that originate nearby, on the
%transverse side will be minimal, as these express low levels of $l_0$. Axons
%originating from the nasal side of the retina also interact most strongly with
%axons originating nearby and in general the $r_0$/$l_0$ interaction is weaker
%between a nasal axon and a second retinal axon. However, when the opposing
%receptor/ligand pair ($r_2$/$l_2$) interaction is summed with that of
%$r_0$/$l_0$, a more even interaction is found, with interactions between axons
%from extreme sides of the retina being the strongest. (Poss.~supp.~figure
%based on the Octave script J.m).

%%\subsection*{Receptor-receptor axon-axon interactions}

\subsection*{Robustness to noise}

\emph{This section might fit most naturally after the Chemoaffinity only section.}

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{./images/model_vs_noise.png}
\caption{ \textbf{A} Chemotaxis-only model. The mean value of $\eta$ (number of crossings) and $\epsilon$
(mean distance error), averaged over the period from simulation time $t=750$
to $t=1000$, plotted with respect to the noise gain ($\nu$) in the best
$\mathbf{G}$ model and the wildtype condition (compare with
Fig.\,\ref{f:G0}). Noise for $\nu < 0.5$ is tolerated by the model, but
disorder increases above this value.
\textbf{B} Competitive interaction in the $\mathbf{GJ}$ model extends the
range of ligand gradient noise which can be tolerated by the system.
Both metrics indicate that a moderate
amount of gradient sampling noise is \emph{beneficial} to the establishment of
an ordered tectum.}
\label{f:noise}
\end{figure}

One problem with axon gradient-following is the ability of the axon to
successfully detect the direction of the gradient given the small distances
over which the axon can sample signalling molecules and the size of receptor
binding noise \citep{goodhill_noise_2014}.  We investigated the ability of the
chemoaffinity-only and the $\mathbf{GJ}$ model systems to self-organise in the
presence of noise, by introducing a stochastic element to each branch's
sampling of the gradient.

% See branch::estimate_ligand_gradient() in sim/agent/branch.h for the code
Noise is added to the estimation of the ligand gradient by multiplying the
true gradient by a random vector, $\mathbf{U}$, whose coordinates are drawn
from a random uniform distribution centred about 0 with magnitude $\nu$,
denoted: $\mathbf{U}(-\nu, \nu)$. Eq.\,\ref{e:G} becomes

\begin{equation}\label{e:Gnu}
\mathbf{G}_b = \sum_i^N F_i\,r_{i,b}\, \mathbf{U}(-\nu, \nu) \nabla L_i(\mathbf{x}_b)
\end{equation}

Fig.\,\ref{f:noise} shows the response of both the chemoaffinity-only model
and the full $\mathbf{GJ}$ model to the addition of ligand gradient estimation
noise. The results indicate that a moderate amount of noise is beneficial for
the establishment of an ordered map and that the competitive model is much
more robust to higher levels of gradient estimation noise.

What about retinal expression noise? This would have a disordering effect on
the pattern.

% What about the EphA manipulations with noise? Ans: No strong effect (just
% add "noise_gain" : 0.5, to one of the expt jsons.

\subsection*{Single axon behaviour}

Acknowledge the result of \citet{gosse_retinotopic_2008} and show the
behaviour of the GC model (and by extension the other models) if a single axon
is allowed to develop.

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{./images/single_axon.png}
\caption{ \textbf{A} Central \textbf{B} Caudal (nasal RGC) \textbf{C} Rostral
(temporal RGC)}
\label{f:single}
\end{figure}

\subsection*{Attractive receptor-ligand interactions}

If we switch a receptor-ligand pair to have an attractive interaction, the
tectal ligand expression should change from exponential to logarithmic.

\section{Discussion}
Discussion goes here.

Discuss the requirement for a distance based weight and the fact that without
this, the system doesn't converge on a stable solution.

% Discuss the Hill Equation! https://en.wikipedia.org/wiki/Hill_equation_(biochemistry)
% This relates to the work \cite{naoki_revisiting_2017}

%
% BIBLIOGRAPHY
%
\selectlanguage{English}
\bibliographystyle{apalike}
\bibliography{RetinoTectal}

%
% Supplementary figures in an appendix
%
\appendix

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\section*{Supplementary figures}

\begin{figure}
\includegraphics[width=\linewidth]{../misc/expression_profiling/expression_profiles.png}
\caption{Expression profiles from the literature.}
\label{f:expressions}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{./images/expressions_manipulations.png}
\caption{Ligand expression and gradients on the tectum for the
other simulated surgical manipulations}
\label{f:tmanip}
\end{figure}

% ./build/sim/agent/agent1 configs/simpler/m_ee_GC_best_1.json configs/simpler/e_wt_figcomp1_exit_true.json
\begin{figure}
\includegraphics[width=\linewidth]{./images/j4_ee_GC_best_1_wt_figcomp1_exit_true.png}
\caption{$\mathbf{GC}$ model}
\label{f:GC}
\end{figure}

% ./scripts/fig_GC_surgical.sh
\begin{figure}
\includegraphics[width=\linewidth]{./images/fig_GC_surgical.png}
\caption{$\mathbf{GC}$ model, surgical manipulations. Demonstrating that this
model performs better than the chemoaffinity-only model.}
\label{f:GCsurg}
\end{figure}

% ./scripts/fig_GClikeGJ_surgical.sh
\begin{figure}
\includegraphics[width=\linewidth]{./images/fig_GClikeGJ_surgical.png}
\caption{$\mathbf{GC}$ model with parameters found from \emph{GJ}
optimization ( $m_G = 0.0012$, $r_c = 0.079$ and $m_c = 0.247$). Results are
equivalent to those for the $\mathbf{GJ}$ model in Fig.\,\ref{f:GJsurg}.}
\label{f:GClikeGJsurg}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{./images/EphA_manipulations_GClikeGJ.png}
\caption{$\mathbf{GC}$ model with parameters found from \emph{GJ}
optimization, genetic manipulations. As for the surgical manipulations, the
results are similar to those of the $\mathbf{GJ}$ model with the equivalent
parameters. However, there are subtle differences such as XXX.}
\label{f:GClikeGJeph}
\end{figure}

% No need to show the GI model
% ./build/sim/agent/agent1 configs/simpler/m_ee_GI_best_1.json configs/simpler/e_wt_figcomp2.json
%\begin{figure}
%\includegraphics[width=\linewidth]{./images/j4_ee_GI_best_1_wt_figcomp2_exit_true.png}
%\caption{$\mathbf{GI}$ model}
%\label{f:GI}
%\end{figure}

% ./scripts/fig_GI_surgical.sh
%\begin{figure}
%\includegraphics[width=\linewidth]{./images/fig_GI_surgical.png}
%\caption{$\mathbf{GI}$ model, surgical manipulations}
%\label{f:GIsurg}
%\end{figure}

\end{document}

%% For the noise based paper:
%% Instead, we treat the receptor binding at the growth cone with greater
%% care. Based on a Michaelis-Menten model applied to
%% receptor binding, \citet{mortimer_bayesian_2009} give an expression for the
%% probability of an individual receptor being bound as a function of the ligand
%% expression, $\gamma$, and the ligand gradient, $\mu$;
%% \begin{equation}
%% P(b=1|\gamma,\mu) = \frac{\gamma(1 + \mu r)}{1 + \gamma(1 + \mu r)},
%% \end{equation}
%% where $r$ is the distance of the receptor from the centre of a 1D growth
%% cone. To determine the growth cone's estimate of the gradient, we assume
%% evenly spaced receptors and sample from a uniform distribution for each
%% location across the growth cone. Activated receptors contribute